---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 3: Portfolio Construction"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Hendri van Zyl"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Stellenbosch, South Africa" # First Author's Affiliation
Email1: "17640296\\@sun.ac.za" # First Author's Email address

# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.
# 
# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"
# 
# CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
# BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
# addtoprule: TRUE
# addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
#Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: Comparison of the SWIX(J403) and ALSI(J203) Indexes
  
---


```{r message=FALSE, warning=FALSE, include=FALSE}
pacman::p_load(tidyverse)
ALSI <- read_rds("/Users/R/Hen_FM_practical/data/ALSI.rds")
RebDays <- read_rds("/Users/R/Hen_FM_practical/data/Rebalance_days.rds")
```


# Introduction {-}

We begin our research into these indexes by first comparing their weighting of constituent stocks. First we will look at sector allocation and then market cap size allocations. 

```{r echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
# Load required libraries
library(dplyr)
library(ggplot2)
library(patchwork) # For combining plots

# Step 1: Aggregate Data by Sector and Date
sector_allocation <- ALSI %>%
  group_by(date, Sector) %>%
  summarise(
    J203_sum = sum(J203, na.rm = TRUE),
    J403_sum = sum(J403, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(date) %>%
  mutate(
    J203_share = J203_sum / sum(J203_sum, na.rm = TRUE),
    J403_share = J403_sum / sum(J403_sum, na.rm = TRUE)
  )

# Step 2: Plot Sector Allocation for J203
plot_j203 <- ggplot(sector_allocation, aes(x = date, y = J203_share, fill = Sector)) +
  geom_area(position = "stack") +
  labs(
    title = "J203",
    y = "Proportion"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none", # legend is shared by both figs
    axis.title.x = element_blank() # Remove x-axis label
  )

# Step 3: Plot Sector Allocation for J403
plot_j403 <- ggplot(sector_allocation, aes(x = date, y = J403_share, fill = Sector)) +
  geom_area(position = "stack") +
  labs(
    title = "J403",
    y = "Proportion"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none", # Legend will be shared
    axis.title.x = element_blank() # Remove x-axis label
  )

# Step 4: Combine the Two Plots with Shared Legend and Main Title
combined_plot <- (plot_j203 / plot_j403) +
  plot_layout(guides = "collect") & 
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

# Add a main title
combined_plot <- combined_plot + plot_annotation(
  title = "Sector Allocation Over Time"
)

combined_plot
#ggsave(here("Question 3/plots/sectweights.png"), plot = combined_plot, width = 8, height = 5, dpi = 300)
```

The sector weight allocations are clearly not exactly the same but they are certainly very close and their reweighting over time seems to rhyme.

```{r echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE}


# Step 1: Update Index_Name for Ticker == "RBW"
ALSI <- ALSI %>%
  mutate(Index_Name = ifelse(Tickers == "RBW", "Mid_Caps", Index_Name))

# Step 2: Aggregate Data by Index_Name and Date
index_allocation <- ALSI %>%
  group_by(date, Index_Name) %>%
  summarise(
    J203_sum = sum(J203, na.rm = TRUE),
    J403_sum = sum(J403, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(date) %>%
  mutate(
    J203_share = J203_sum / sum(J203_sum, na.rm = TRUE),
    J403_share = J403_sum / sum(J403_sum, na.rm = TRUE)
  )

# Step 3: Plot Allocation for J203
plot_j203 <- ggplot(index_allocation, aes(x = date, y = J203_share, fill = Index_Name)) +
  geom_area(position = "stack") +
  labs(
    title = "J203",
    y = "Proportion"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none", # Legend will be shared
    axis.title.x = element_blank() # Remove x-axis label
  )

# Step 4: Plot Allocation for J403
plot_j403 <- ggplot(index_allocation, aes(x = date, y = J403_share, fill = Index_Name)) +
  geom_area(position = "stack") +
  labs(
    title = "J403",
    y = "Proportion"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none", # Legend will be shared
    axis.title.x = element_blank() # Remove x-axis label
  )

# Step 5: Combine the Two Plots with Shared Legend and Main Title
combined_plot <- (plot_j203 / plot_j403) +
  plot_layout(guides = "collect") & 
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

# Add a main title
index_plot <- combined_plot + plot_annotation(
  title = "Index Allocation Over Time"
)

# Display the Combined Plot
index_plot

# ggsave(here("Question 3/plots/index_plot.png"), plot = index_plot, width = 8, height = 5, dpi = 300)
```

Index allocations are also very similar. A surprising amount of NAs. Both greatly favour large caps while mid caps make up about a quarter of the allocation and small caps are a trivial component. J403 appears to have a slightly higher allocation in mid caps while J203 appear to have a relatively higher allocation in large caps, this is especially in the most recent months. 

# Performance

```{r echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
# Load required libraries
library(dplyr)
library(ggplot2)

# Step 1: Calculate Weighted Returns for Each Index
cumulative_returns <- ALSI %>%
  group_by(date) %>%
  summarise(
    J203_weighted_return = sum(Return * J203, na.rm = TRUE), # Weighted return for J203
    J403_weighted_return = sum(Return * J403, na.rm = TRUE), # Weighted return for J403
    .groups = "drop"
  ) %>%
  mutate(
    J203_cum_return = cumprod(1 + J203_weighted_return), # Cumulative return for J203
    J403_cum_return = cumprod(1 + J403_weighted_return)  # Cumulative return for J403
  )

# Step 2: Reshape Data for Plotting
cumulative_returns_long <- cumulative_returns %>%
  select(date, J203_cum_return, J403_cum_return) %>%
  pivot_longer(
    cols = starts_with("J"),
    names_to = "Index",
    values_to = "Cumulative_Return"
  )

# Step 3: Plot Cumulative Returns
cumulative_plot <- ggplot(cumulative_returns_long, aes(x = date, y = Cumulative_Return, color = Index)) +
  geom_line(size = 1) +
  labs(
    title = "Cumulative Returns for J203 and J403",
    x = "",
    y = "Cumulative Return",
    color = "Index"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
cumret <- fmxdat::finplot(cumulative_plot, x.date.dist = "1 year", x.date.type = "%Y", x.vert = T,
    y.pct = T, y.pct_acc = 1)

# Display the Plot
print(cumret)
#ggsave(here("Question 3/plots/cumret.png"), plot = cumret, width = 8, height = 5, dpi = 300)

```

Cumulative returns for the indexes closely tracked each other up until 2020 after which they diverged and with J203 pulling ahead. This might be related to J203's heavier allocation in large caps relative to J403. 


Below we compare rolling standard deviation as a proxy for volatility and risk. 

```{r echo=FALSE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
library(RcppRoll)

# Step 1: Calculate weighted returns for each index
rolling_sd_df <- ALSI %>%
  group_by(date) %>%
  summarise(
    J203_weighted_return = sum(Return * J203, na.rm = TRUE), # Weighted return for J203
    J403_weighted_return = sum(Return * J403, na.rm = TRUE), # Weighted return for J403
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = starts_with("J"),
    names_to = "Index",
    values_to = "Returns"
  ) %>%
  group_by(Index) %>%
  mutate(
    RollSD = RcppRoll::roll_sd(1 + Returns, 24, fill = NA, align = "right") * sqrt(12) # Annualized rolling SD
  ) %>%
  filter(!is.na(RollSD))

# Step 2: Plot the rolling 2-year standard deviation
i <- rolling_sd_df %>%
  ggplot() +
  geom_line(aes(date, RollSD, color = Index), alpha = 0.7, size = 1.25) +
  labs(
    title = "Rolling 2-Year Annualized Standard Deviation Comparison",
    x = "",
    y = "Rolling 2-Year Returns (Ann.)",
    caption = "Note: Monthly data, annualized SD calculated over a 24-month rolling window"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    legend.position = "bottom"
  )

# Display the plot
print(i)
```

The risk profile according to the standard deviation measure is remarkably similar especially given the divergence in cumulative returns in the last two years. 


The results of the expected shortfall comparison between the indexes support this finding. 
```{r message=FALSE, warning=FALSE, include=FALSE}
# Load required libraries
library(dplyr)
library(PerformanceAnalytics)

# Step 1: Filter data for the last 2 years
last_two_years <- ALSI %>%
  filter(date >= max(date) - lubridate::years(2)) %>%
  group_by(date) %>%
  summarise(
    J203_weighted_return = sum(Return * J203, na.rm = TRUE),
    J403_weighted_return = sum(Return * J403, na.rm = TRUE),
    .groups = "drop"
  )

# Step 2: Compute Expected Shortfall
# Convert data to xts for PerformanceAnalytics compatibility
returns_xts <- xts::xts(
  last_two_years %>% select(J203_weighted_return, J403_weighted_return),
  order.by = last_two_years$date
)

# Calculate Expected Shortfall (at 95% confidence level)
es_values <- PerformanceAnalytics::ES(
  R = returns_xts,
  p = 0.05,
  method = "historical"
)

# Step 3: Print the Expected Shortfall results
es_results <- tibble::tibble(
  Index = c("J203", "J403"),
  Expected_Shortfall = as.numeric(es_values)
)

# Display the results
print(es_results)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create gt Table
library(gt)
es_results %>%
  gt() %>%
  tab_header(
    title = "Expected Shortfall Results"
  ) %>%
  fmt_number(
    columns = vars(Expected_Shortfall),
    decimals = 4
  )
```


J203 only displays a marginally more negative expected shortfall than does J403 which indicates that both have similar levels of tail risk in extreme conditions. 



