---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 2: Currency Hedging analysis"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Hendri van Zyl"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Stellenbosch, South Africa" # First Author's Affiliation
Email1: "17640296\\@sun.ac.za" # First Author's Email address

# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.
# 
# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"
# 
# CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
# BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
# addtoprule: TRUE
# addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
#Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: This paper investigates the volatility paradox phenomenon whereby negatively correlated assets in a portfolio counterintuitively result in lower overall volatility than would be expected given the constituent assets. 
  
---


# The Volatility paradox

The figure below illustrates that global assets and the Rand have a demonstrably negative relationship over the sample period. 

```{r include=FALSE}
library(tidyverse)
library(Texevier)
wd <- "/Users/R/Hen_FM_practical/"
Indexes <- read_rds("/Users/R/Hen_FM_practical/data/Cncy_Hedge_Assets.rds")
ZAR <- read_rds("/Users/R/Hen_FM_practical/data/Monthly_zar.rds")
```


```{r message=FALSE, warning=FALSE, include=FALSE}
# Merge ZAR exchange rate with Indexes
data <- Indexes %>%
  mutate(YM = format(date, "%Y_%b")) %>% 
  left_join(ZAR %>% 
              mutate(ZAR_Return = value/lag(value)-1)%>% 
              mutate(YM = format(date, "%Y_%b")) %>% 
              select(-c(date,Tickers)), by = "YM") 



## Convert USD returns to ZAR returns
#data_ZAR <- data %>%
#  mutate(
#    MSCI_ACWI_ZAR = MSCI_ACWI * value,  # Convert MSCI_ACWI to ZAR
#    Bbg_Agg_ZAR = Bbg_Agg * value)  # Convert BbgAfg to ZAR

# Convert ZAR to USD
data_USD <- data %>% 
  arrange(date) %>% 
  group_by(date) %>% 
  mutate(ALBI_USD = ALBI/value,
         J433_USD = J433/value) %>% 
  ungroup()

## Calculate portfolio returns
#Portf <- data_ZAR %>%
#  mutate(
#    # Hedged portfolio (fully in ZAR)
#    portfolio_hedged = 0.42 * J433 + 0.28 * ALBI + 0.12 * Bbg_Agg_ZAR + 0.18 * MSCI_ACWI_ZAR,
#    
#    # Unhedged portfolio (global components remain in USD)
#    portfolio_unhedged = 0.42 * J433 + 0.28 * ALBI +
#      0.12 * Bbg_Agg + 0.18 * MSCI_ACWI
#  )

Portf <- data_USD %>% 
  group_by(date) %>% 
  mutate(
    # Hedged portfolio (fully in USD)
    portfolio_unhedged = 0.42 * J433_USD + 0.28 * ALBI_USD + 0.12 * Bbg_Agg + 0.18 * MSCI_ACWI,
    
    # Unhedged portfolio (global components remain in USD)
    portfolio_hedged = 0.42 * J433 + 0.28 * ALBI +0.12 * Bbg_Agg + 0.18 * MSCI_ACWI
  ) %>% 
  ungroup()
  

```

```{r echo=FALSE, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
# Load necessary libraries
library(ggplot2)
library(ggExtra)
library(fmxdat)

# Calculate quadrant percentages
Portf <- Portf %>%
  mutate(
    quadrant = case_when(
      ZAR_Return < 0 & portfolio_hedged >= 0 ~ "Top Left",
      ZAR_Return >= 0 & portfolio_hedged >= 0 ~ "Top Right",
      ZAR_Return < 0 & portfolio_hedged < 0 ~ "Bottom Left",
      ZAR_Return >= 0 & portfolio_hedged < 0 ~ "Bottom Right"
    ),
    restricted_quadrant = case_when(
      ZAR_Return*100 < -2.5 & portfolio_hedged >= 0 ~ "Restricted Top Left",
      ZAR_Return*100 < -2.5 & portfolio_hedged < 0 ~ "Restricted Bottom Left",
      ZAR_Return*100 >= -2.5 & portfolio_hedged >= 0 ~ "Restricted Top Right",
      ZAR_Return*100 >= -2.5 & portfolio_hedged < 0 ~ "Restricted Bottom Right",
      TRUE ~ NA_character_
    )
  )

# Calculate percentages for original quadrants
quadrant_counts <- Portf %>%
  group_by(quadrant) %>%
  summarise(n = n()) %>%
  mutate(percentage = round(n / sum(n) * 100, 1))

# Calculate percentages for restricted quadrants
restricted_counts <- Portf %>%
  filter(!is.na(restricted_quadrant)) %>%  # Only include restricted quadrants
  group_by(restricted_quadrant) %>%
  summarise(n = n()) %>%
  mutate(percentage = round(n / sum(n) * 100, 1))

# Prepare percentages for labels
percentage_TopLeft <- paste0(quadrant_counts$percentage[quadrant_counts$quadrant == "Top Left"], "%")
percentage_TopRight <- paste0(quadrant_counts$percentage[quadrant_counts$quadrant == "Top Right"], "%")
percentage_BottomLeft <- paste0(quadrant_counts$percentage[quadrant_counts$quadrant == "Bottom Left"], "%")
percentage_BottomRight <- paste0(quadrant_counts$percentage[quadrant_counts$quadrant == "Bottom Right"], "%")

# Restricted quadrant percentages
percentage_RestrictedTopLeft <- paste0(
  restricted_counts$percentage[restricted_counts$restricted_quadrant == "Restricted Top Left"], "%"
)
percentage_RestrictedBottomLeft <- paste0(
  restricted_counts$percentage[restricted_counts$restricted_quadrant == "Restricted Bottom Left"], "%"
)

# Generate scatter plot with updated quadrant label positions
scatter_plot <- ggplot(Portf, aes(x = ZAR_Return * 100, y = portfolio_hedged * 100)) +
  # Add pastel background quadrants with gridlines visible
  geom_rect(aes(xmin = -20, xmax = 0, ymin = 0, ymax = 20), fill = "#FFDAB9", alpha = 0.9) +  # Light peach
  geom_rect(aes(xmin = 0, xmax = 20, ymin = 0, ymax = 20), fill = "#E6E6FA", alpha = 0.9) +  # Lavender
  geom_rect(aes(xmin = -20, xmax = 0, ymin = -20, ymax = 0), fill = "#B0E0E6", alpha = 0.5) + # Powder blue
  geom_rect(aes(xmin = 0, xmax = 20, ymin = -20, ymax = 0), fill = "#98FB98", alpha = 0.5) +  # Pale green

  # Add scatter points
  geom_point(alpha = 0.6) +

  # Add density contours with thinner and more opaque lines
  geom_density2d(color = "blue", size = 0.3, alpha = 0.8) +

  # Add regression line
  geom_smooth(method = "lm", color = "red", linetype = "solid", alpha = 0.9) +

  # Add solid zero lines
  geom_hline(yintercept = 0, linetype = "solid", color = "grey") +
  geom_vline(xintercept = 0, linetype = "solid", color = "grey") +

  # Add dashed line for hedging fee at -2.5%
  geom_vline(xintercept = -2.5, linetype = "dashed", color = "grey") +

  # Add quadrant percentage labels
  geom_label(aes(x = -20, y = 20, label = percentage_TopLeft), hjust = 0, vjust = 1, 
             size = 4, color = "black", fill = "white") +  # Top-left quadrant
  geom_label(aes(x = 20, y = 20, label = percentage_TopRight), hjust = 1, vjust = 1, 
             size = 4, color = "black", fill = "white") +  # Top-right quadrant (moved to top-right corner)
  geom_label(aes(x = -20, y = -20, label = percentage_BottomLeft), hjust = 0, vjust = 0, 
             size = 4, color = "black", fill = "white") + # Bottom-left quadrant
  geom_label(aes(x = 20, y = -20, label = percentage_BottomRight), hjust = 1, vjust = 0, 
             size = 4, color = "black", fill = "white") + # Bottom-right quadrant (moved to bottom-right corner)

  # Add restricted quadrant percentage labels
  geom_label(aes(x = -2.5, y = 20, label = percentage_RestrictedTopLeft), hjust = 1, vjust = 1, 
             size = 4, color = "black", fill = "white") +  # Restricted top-left quadrant
  geom_label(aes(x = -2.5, y = -20, label = percentage_RestrictedBottomLeft), hjust = 1, vjust = 0, 
             size = 4, color = "black", fill = "white") + # Restricted bottom-left quadrant

  # Add quadrant annotations with geom_label()
  geom_label(aes(x = -10, y = 13, label = "Hedge works but \n amplifies volatility"),
             hjust = 0.5, vjust = 0.5, size = 4, color = "black", fill = "white", alpha = 0.8) +
  geom_label(aes(x = 10, y = 13, label = "Hedge throws away \n returns"),
             hjust = 0.5, vjust = 0.5, size = 4, color = "black", fill = "white", alpha = 0.8) +
  geom_label(aes(x = -10, y = -13, label = "Best case for hedge: \n higher return, lower volatility"),
             hjust = 0.5, vjust = 0.5, size = 4, color = "black", fill = "white", alpha = 0.8) +
  geom_label(aes(x = 10, y = -13, label = "Hedge removes \n currency cushion"),
             hjust = 0.5, vjust = 0.5, size = 4, color = "black", fill = "white", alpha = 0.8) +

  # Set axis limits and labels
  scale_x_continuous(limits = c(-20, 20), breaks = seq(-20, 20, 10), labels = function(x) paste0(x, "%")) +
  scale_y_continuous(limits = c(-20, 20), breaks = seq(-20, 20, 10), labels = function(y) paste0(y, "%")) +

  # Add labels and theme
  labs(
    title = "Scatter Plot of USD/ZAR returns vs Portfolio returns",
    x = "USD-ZAR Exchange Rate Returns (%)",
    y = "60/40 Global Portfolio Returns (USD) (%)"
  ) +
  theme_minimal()

# Add marginal density plots with increased opacity
ggExtra::ggMarginal(scatter_plot, type = "density", fill = "grey", alpha = 0.7)
```


# Rolling volatity

The figure below clearly illustrates that using the rand as a currency hedge increases volatility substantially. 

```{r echo=FALSE, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
combo_df <- Portf %>% 
  select(date, portfolio_hedged, portfolio_unhedged) %>% 
  pivot_longer(-date, names_to = "Type", values_to = "Returns") %>% 
  group_by(Type) %>% 
  mutate(RollSD = RcppRoll::roll_sd(1 + Returns, 24, fill = NA, align = "right") *
    sqrt(12)) %>%
    filter(!is.na(RollSD))

i <- combo_df %>%
    ggplot() + geom_line(aes(date, RollSD, color = Type),
    alpha = 0.7, size = 1.25) + labs(title = "Rolling 2 Year Annualized Standard Devation Comparison",
    subtitle = "", x = "", y = "Rolling 2 year Returns (Ann.)",
    caption = "Note:") + theme_fmx(title.size = ggpts(30),
    subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = T) +
    fmx_cols()

roll_sd_port <- finplot(i, x.date.dist = "1 year", x.date.type = "%Y", x.vert = T,
    y.pct = T, y.pct_acc = 1)

#ggsave(here("Question 2/plots/roll_sd_port.png"), plot = roll_sd_port, width = 6, height = 4, dpi = 300)

roll_sd_port
```


This strengthens the argument against long term currency hedging for the purposes of reducing overall volatility by introducing a negatively correlated asset. One can at least say that in this very specific context the strategy does not appear to be effective. 




